---
title: Basketball Data Sciene Basics
author: Randi Bolt
date: '2022-04-22'
slug: []
categories:
  - undergrad
tags: []
---

Recently I started reading [Bsketball Data Sciene]().... here are some things. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(error=FALSE,
                      message= FALSE,
                      warning=FALSE)
```

# Load Packages 

Using .. 

```{r}
library(BasketballAnalyzeR)
library(ggplot2)
```


# Data

```{r}
BAR <- data(package = "BasketballAnalyzeR")
head(BAR)
```

There are 5 data sets:

1. Obox: Opponents box scores dataset - NBA 2017-2018

2. Pbp.BDB: Play by Play dataset - NBA 2017-2018

3. Pbox: Players box scores dataset - NBA 2017-2018

4. Tadd: Team information such as Conference, Division, final rank, qualification for Playoffs for the NBA 2017-2018 Championship.

5. Tbox: Team box scores dataset - NBA 2017-2018

```{r}
Pbp <- PbPmanipulation(PbP.BDB)
head(Pbp)
```


# Concepts (MATH)

### Possession and Pace

$$\text{POSS = (P2A + P3A) + 0.44 }\times \text{ FTA - OREB + TOV}$$ 

$$\text{PACE = 5 }\times\text{ POSS/MIN}$$  

Where MIN is the total minutes played by all the players. 

### Offensive and Defensive Ratings

Pre-possession efficiency is measure as the points scored or allowed per 100 possession, called Offensive (ORtg) and Defensive (DRtg) Ratings respectively: 

$$\text{ORtg = PTS}_\text{T}/\text{POSS}_{\text{T}}$$

$$\text{DRtg = PTS}_\text{O}/\text{POSS}_{\text{O}}$$

Where T and O refer to team (being analyzed) and opponent(s). 

### Four Factors

$\underline{\text{Efficient Field Goal Percentage}}$ 

$$\text{eFG }\%_{\text{ Offense}}=\frac{\text{P2M}_\text{T}+1.5\times\text{ P3M}_\text{T}}{\text{P2A}_\text{T}+\text{P3A}_\text{T}}$$

$$\text{eFG }\%_\text{ Defense}=\frac{\text{P2M}_\text{O}+1.5\times\text{ P3M}_\text{O}}{\text{P2A}_\text{O}+\text{P3A}_\text{O}}$$

$\underline{\text{Turnovers Per Possession Ratio}}$

$$\text{TO Ratio}_\text{ Offense}=\frac{\text{TOV}_\text{T}}{\text{POSS}_\text{T}}$$

$$\text{TO Ratio}_\text{ Defense}=\frac{\text{TOV}_\text{O}}{\text{POSS}_\text{O}}$$

$\underline{\text{Rebounding Percentage}}$

$$\text{REB }\%_{\text{ Offense}}=$$

$$\text{REB }\%_\text{ Defense}=$$

$\underline{\text{Free Throw Rate}}$

$$\text{FT Rate}_{\text{ Offense}}=$$

$$\text{FT Rate}_{\text{ Defense}}=$$

# Plots

```{r}
tm <- c("BOS", "CLE", "GSW", "HOU")
selTeams <- which(Tadd$team %in% tm)
FF.sel <- fourfactors(Tbox[selTeams,], Obox[selTeams,])
plot(FF.sel)
```

```{r}
# rearrange 
listPlots <- plot(FF.sel)
library(gridExtra)
grid.arrange(grobs=listPlots[1:2], ncol=1)
grid.arrange(grobs=listPlots[3:4], ncol=1)
```

### Bar line Plots

```{r}
X <- data.frame(Tbox, 
                PTS.O = Obox$PTS, 
                TOV.O = Obox$TOV, 
                CONF = Tadd$Conference)
XW <- subset(X, CONF== "W")
labs <- c("Steals", "Blocks", "Defensive Rebounds")
barline(data = XW,
        id = "Team",
        bars = c("STL", "BLK", "DREB"),
        line = "TOV.O",
        order.by = "PTS.O",
        labels.bars = labs)
```

```{r}
Pbox.HR <- subset(Pbox, 
                  Team=="Phoenix Suns" & MIN>= 500)
Pbox.HR
barline(data = Pbox.HR, id = "Player",
        bars = c("P2p", "P3p", "FTp"),
        line = "MIN",
        order.by = "PM",
        labels.bars = c("2P%", "3P%", "FT%"),
        title = "Phoenix Suns")
```

### Radial Plots

```{r}
Pbox.PG <- subset(Pbox, 
                  Player == "Russell Westbrook" |
                    Player == "Stephen Curry" |
                    Player == "Kyrie Irving" |
                    Player == "Damian Lillard" |
                    Player == "Kyle Lowry" |
                    Player == "John Wall" |
                    Player == "Rajon Rondo" |
                    Player == "Kemba Walker")
attach(Pbox.PG)
X <- data.frame(P2M, P3M, FTM, REB=OREB+DREB, AST, STL, BLK)/MIN
detach(Pbox.PG)
radialprofile(data = X, title = Pbox.PG$Player, std = F)
radialprofile(data = X, title = Pbox.PG$Player, std = T)
```

### Scatter Plots

```{r}
Pbox.sel <- subset(Pbox, MIN>=500)
attach(Pbox.sel)
X <- data.frame(AST, TOV, PTS)/MIN
detach(Pbox.sel)
mypal <- colorRampPalette(c("blue", "yellow", "red"))
scatterplot(X, data.var = c("AST", "TOV"), z.var="PTS", labels = 1:nrow(X), palette = mypal)
SAS <- which(Pbox.sel$Team=="San Antonio Spurs")
scatterplot(X, data.var = c("AST", "TOV"), z.var="PTS", labels = Pbox.sel$Player, palette = mypal, subset = SAS)
scatterplot(X, data.var = c("AST", "TOV"), z.var="PTS", labels = Pbox.sel$Player, palette = mypal, subset = SAS, zoom = c(0.2, 0.325, 0.05, 0.10))
```

### Bubble Plots 

```{r}
attach(Tbox)
X <- data.frame(T = Team, P2p, P3p, FTp, AS = P2A + P3A + FTA)
detach(Tbox)
labs <- c("2-Point shots (% made)",
          "3-Point shots (% made)",
          "Free throws (% made)",
          "Total shots attempted")
bubbleplot(X, id = "T", x = "P2p", y = "P3p", col = "FTp", 
           size = "AS", labels = labs)
```


```{r}
Pbox.CC.GSW <- subset(Pbox, 
                       Team == "Golden State Warriors" | Team == "Cleveland Cavaliers" & MIN >= 500)
attach(Pbox.CC.GSW)
X <- data.frame(ID = Player,
                Team,
                V1 = DREB/MIN,
                V2 = STL/MIN,
                V3 = BLK/MIN,
                V4 = MIN)
detach(Pbox.CC.GSW)
labs <- c("Defensive Rebounds",
          "Steals",
          "Blocks",
          "Total minutes played")
bubbleplot(X, id = "ID",
           x = "V1", y = "V2", col = "V3", size = "V4",
           text.col = "Team", labels = labs, 
           title = "CC and GSW during the regular season",
           text.legend = TRUE,
           text.size = 3.5,
           scale = FALSE)
```

# Varibility Analysis


