---
title: reactable
author: 'Randi Bolt '
date: '2022-03-24'
slug: []
categories:
  - undergrad
tags:
  - react
  - rvest
---

# Background 

I wanted to try reactable, and NBA playoffs start in April. 

# Set Up 

Handy dandy options, 

```{r options}
# Global Options 
knitr::opts_chunk$set(error=FALSE,
                      message= FALSE,
                      warning=FALSE)
```

and libraries. 

```{r set up}
# Load Libraries 
library(rvest) 
library(tidyverse) 
library(reactable)
```

# Harvest the data with rvest

I could easily pull the CSV's off [Basketball Reference](https://www.basketball-reference.com/), but I might want to come back and adjust this post later. It would be pretty tedious to save those data sets again, so instead I am using the package `rvest` to harvest the data directly off the web page. For this table I am pulling the top team in the west, which is the Phoenix Suns and the top teams in the east, which is the Miami Heat.   

```{r variables}
# Add team name and team slug
suns <- "Phoenix Suns"
suns_slug <- "PHO"

heat <- "Miami Heat"
heat_slug <- "MIA"

# define player page URL
suns_url <- paste0("https://www.basketball-reference.com/teams/",suns_slug,"/2022.html")

heat_url <- paste0("https://www.basketball-reference.com/teams/",heat_slug,"/2022.html")
```

If you check either of these teams Basketball Reference page you will see there is a lot of data and statistics available. Since I really just want to try reactable, lets just look at the Total stats. 

```{r rvest data}
# Read total stats
suns_ttl_stat <- suns_url %>%
  read_html %>%
  html_node("#totals") %>% 
  html_table()

heat_ttl_stat <- heat_url %>%
  read_html %>%
  html_node("#totals") %>% 
  html_table()
```

# Tidy 

To tidy this data there are four main things I am doing: 

- add "Name" to the column with players names

- add column with team name (as to be able to sort by team)

- merge Suns and Heats tables

- Remove NA rows and extra columns 

```{r tidy data}
# Rename Column 2 Name 
names(suns_ttl_stat)[2] <- "Name"
names(heat_ttl_stat)[2] <- "Name"

# Add Column with Team Name
suns_ttl_stat$Team <- suns
heat_ttl_stat$Team <- heat

# Merge stats tables
total_stats <- full_join(suns_ttl_stat, heat_ttl_stat) 

# Remove NA rows and extra column data 
total_stats <- total_stats %>% 
  filter(!is.na(Rk)) %>%
  select(Team, Name, Rk, Age, G, MP, PTS, FG, `3P`, `2P`, FT, AST, STL, BLK, TOV, PF)

# View Data
head(total_stats)
```

# Reactable 

There were five main things I wanted from this table: 

- groupBy = Team 

- remane columns

- bordered

- highlight

- searchable

- minRows = number of teams showing (which is 2)

```{r reactable}
reactable(
  total_stats,
  groupBy = "Team",
  columns = list(
    Rk = colDef(name = "Rank"),
    G = colDef(name = "Games"),
    MP = colDef(name = "Minutes Played"),
    FG = colDef(name = "Field Goals"),
    `3P` = colDef(name = "3 Point Goals"),
    `2P` = colDef(name = "2 Point Goals"),
    FT = colDef(name = "Free Throws"),
    AST = colDef(name = "Assists"),
    STL = colDef(name = "Steals"),
    BLK = colDef(name = "Blocks"),
    PTS = colDef(name = "Points"),
    TOV = colDef(name = "Turnovers"),
    PF = colDef(name = "Personal Fouls")
  ),
  bordered = TRUE,
  highlight = TRUE,
  filterable = TRUE,
  searchable = TRUE,
  minRows = 2
  )
```

# Sources

[Reactable](https://glin.github.io/reactable/index.html)

[NBA Analytics Tutorial: Using R to Display Player Career Stats](https://sweepsportsanalytics.com/2021/12/19/nba-analytics-tutorial-using-r-to-display-player-career-stats/)

# Future Updates

- Create a question to look at

- Add Functions to look at more teams / reduce code

- Add Graphs or other visuals 

Clean Up Reactable:

- Add logos 

- Add Graphs 

- Narrow Data and Compute Statistics

- Add more teams 

