---
title: 'Bayes Rules! - Salt Lake City R User Group '
author: Randi Bolt
date: '2022-04-06'
slug: []
categories:
  - undergrad
tags: []
---

 Next Meeting: https://www.meetup.com/slc-rug/events/284188359/ 
  
  Code of Conduct: https://wiki.r-consortium.org/view/R_Consortium_and_the_R_Community_Code_of_Conduct 

If you are interested in speaking, reach out. 

Alicia Johnson (speaker)

- statistics professor 

- author of Bayes Rules!
  
# Links
  
Workshop Materials: https://docs.google.com/presentation/d/13SnjPZsJpJWBHimIhA2edBoQHIp0_1xkF-k2T8uj2pg/edit#slide=id.p

Github Repository: https://github.com/ajohns24/r-ladies-philly

Website: https://www.bayesrulesbook.com/ 
  
# Quiz 
  
**1. What does P(heads)=0.5 mean?**
  
    a. If I flip this coin over and over, roughly 50% will be heads. 

    b. Heads and Tails are equally plausible.

    c. Both a and b make sense. 

>Majority Responses: C
>
>Scores: a = 1, b = 3, c = 2 

a) Frequentist Philosophy 

    - long run outcome 

b) Bayesian Philosophy

    - relative probability of events 

c) Pragmatic Philosophy 

    - both interpretations make sense

**2. P(candidate A wins) = 0.8 means**

    a. If we observe this election over and over, candidate A will win roughly 80% of the time. 

    b. Candidate A is much more likely to win than to lose (4 times more likely).

    c. The pollster's calculation is wrong. 

>Majority Response: B 
>
>Scores: a = 1, b = 3, c = 1

a) Freq. 

b) Bayes 

c) Freq. 

    - b/c event cannot be repeated over and over again 

**3. Alicia claims she can predict the outcome of a coin flip.** 

**Mine claims she can distinguish between a Crown Burger and a vegan alternative.** 

**Both succeed in 10 out of 10 trials! What do you conclude?** 

    a. You're still more confident in Mine's claim than Alicia's claim.

    b. The evidence supporting Mine's claim. 

>Score: a = 3, b = 1

a) Bayes

b) Freq. 

**4. You've tested positive for a very rare genetic trait. If you only get to ask the doctor one question, which would it be?**
  
    a. P(rare trait|+)

    b. P(+| no rare trait)

> Score: a = 3, b = 1 

a) Bayes

    - asking about uncertainty of hypothesis given certainty of the data

    - more natural question to ask 

b) Freq. = p-value 

    - hard to wrap minds around 

    - asking about uncertainty in data 

    - less natural question to ask (since data is certain)

# Bayes Rules Activity 

```{r echo = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  fig.height = 2.5,
  fig.width = 4.5,
  fig.pos = "center")
```


## Goals 

- Learn to think like Bayesians.
- Apply Bayesian thinking in a regression setting.

## Set Up

```{r}
# Load packages
library(tidyverse)
library(tidybayes)
library(bayesrules)
library(bayesplot)
library(rstanarm)
library(broom.mixed)
```

## Background 

Let $\pi$ ("pi") be the proportion of U.S. adults that believe that climate change is real and caused by people. Thus $\pi$ is some value between 0 and 1.

### Exercise 1: Specify a prior model

The first step in learning about $\pi$ is to specify a **prior model** for $\pi$ (i.e. prior to collecting any data). Suppose your friend specifies their understanding of $\pi$ through the "Beta(2, 20)" model. Plot this Beta model and discuss what it tells you about your friend's prior understanding. For example:

- What do they think is the most likely value of $\pi$?    
- What range of $\pi$ values do they think are plausible?

```{r}
plot_beta(alpha = 2, beta = 20)
```

Notes:

- proportion between 0 and 1 (not $-\infty$ to $+\infty$)

- this model, beta-2-20, is right skewed 

Q: what is your friend saying is the most likely value of pie?

A: about .12, aka about 12% of people believe in climate change. 

- spike of model is best estimate 

- looking at range the prior model drops off above .25, so you friend believes under 25% of people believe in climate change. 

### Exercise 2: Check out some data

The second step in learning about $\pi$, the proportion of U.S. adults that believe that climate change is real and caused by people, is to collect data. Your friend surveys 10 people and 6 believe that climate change is real and caused by people. The **likelihood function** of $\pi$ plots the chance of getting this 6-out-of-10 survey result under different possible $\pi$ values. Based on this plot:

- With what values of $\pi$ are the 6-out-of-10 results most consistent?

    Approx. 60% of people belive in climate change. Shown by our graph spiking at that value. 

- For what values of $\pi$ would these 6-out-of-10 results be *unlikely*?

    Our data would not be very likely to happen for values below .25 and above .9. 

```{r}
plot_binomial_likelihood(y = 6, n = 10)
```

Notes: 

- The next step after creating a model (beta-2-20) we collect data. 

- This plot is showing us what the chance is that we got these survey results under different possible pie values. 

### Exercise 3: Build the posterior model

In a Bayesian analysis of $\pi$, we build a *posterior* model of $\pi$ by combining the prior model of $\pi$ with the data (represented through the likelihood function). Plot all 3 components below. Summarize your observations:

- What's your friend's posterior understanding of $\pi$?

    My friend's prior understanding of $\pi$ is not as low as what it was before, but also not as high as what is suggested in the data. 

- How does their posterior understanding compare to their prior and likelihood? Thus how does their posterior *balance* the prior and data?


```{r}
plot_beta_binomial(alpha = 2, beta = 20, y = 6, n = 10)
```

Notes:

- Depends on a lot of factors 

### Exercise 4: Another friend

Consider another friend that saw the same 6-out-of-10 polling data but started with a Beta(1, 1) prior model for $\pi$:

```{r}
plot_beta(alpha = 1, beta = 1)
```

- Describe the new friend's understanding of $\pi$. Compared to the first friend, are they more or less sure about $\pi$?

- Do you think the new friend will have a different posterior model than the first friend? If so, how do you think it will compare?

- Test your intuition. Use `plot_beta_binomial()` to explore your new friend's posterior model of $\pi$.

```{r}
plot_beta_binomial(alpha = 1, beta = 1, y = 6, n = 10)
```

Notes:

- this is a shoulder shrug, uncertain prior model. It could really be anything. 

### Exercise 5: More data

Your two friends come across more data. In the `pulse_of_the_nation` survey, 655 of 1000 people believed climate change was real and caused by people:

```{r}
data("pulse_of_the_nation")
pulse_of_the_nation %>% 
  count(climate_change)
```

- How do you think the additional data will impact your first friend's posterior understanding of $\pi$? What about the second friend's?

- Upon seeing the 1000-person survey results, do you think your two friends' posterior understandings of $\pi$ will disagree a lot or a little?

- Test your intuition. Use `plot_beta_binomial()` to explore both friends' posterior models of $\pi$.

```{r}
# first friend 
plot_beta_binomial(alpha = 2, beta = 20, y = 655, n = 1000)
# second friend 
plot_beta_binomial(alpha = 1, beta = 1, y = 655, n = 1000)
```

--> See the link to the github repository above for the rest of this activity.  
