---
title: 'Reproducible Circular NBA Barcharts '
author: Randi Bolt
date: '2022-06-02'
slug: []
categories:
  - Project Engineering
tags:
  - R
  - docker
  - Makefile
  - NBAcharts
  - rvest
---

These are the steps I took to re-create NBA Circular Barcharts, but in a structured reproducible manner with a Makefile, and Dockerfile. You can find the _completed_ version of this project  [here](https://github.com/rbolt13/nba_proj). 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# libraries 
library(kableExtra)
library(magrittr)
# function that creates tables 
tbs_function <- function(terminal, description){
  df <- base::data.frame("Terminal_Command" = terminal, 
                 "Description" = description)
 kb_df <- kableExtra::kable(df,
                  col.names = c("Terminal Command", 
                                "Description"),
                  align = "cc") %>%
  kable_material_dark("hover", full_width = F) 
 return(kb_df)
}
```

# 00_ From the beginning

<center>

<iframe src="https://giphy.com/embed/Vh2szshWnZz9D7AHDB" width="480" height="270" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p><a href="https://giphy.com/gifs/cartoonnetwork-steven-universe-the-movie-sutm-Vh2szshWnZz9D7AHDB">via GIPHY</a></p>

</center>

### 00_00_ Folders

Open up a bash terminal and type in the following lines of code:  

```{r creating project folders, echo=FALSE}
terminal <- c("mkdir nba_proj",
              "cd nba_proj",
              "pwd",
              "mkdir 00_R",
              "mkdir 01_raw_data",
              "mkdir 02_clean_data",
              "mkdir 03_figs",
              "mkdir 04_output",
              "mkdir 05_sandbox")
description <- c("make nba_proj folder",
                 "move into nba_proj folder",
                 "print working directory",
                 "make 00_R folder for the code",
                 "make 01_raw_data folder for the raw (scraped) data",
                 "make 02_clean_data folder for the clean data",
                 "make 03_figs folder for the barchart imgs",
                 "make 04_output folder for report that is printed",
                 "make 05_sandbox folder as a space to play with the code")
tbs_function(terminal, description)
```

Note: I numbered my folders so they would be in the general _order_ I would use them / would want a collaborator to view them. 

### 00_01_ renv

```{r renv, echo=FALSE}
terminal <- c("R",
              "ren::init()",
              "q()",
              "y")
description <- c("open R within the terminal",
                 "creates renv folder and contents",
                 "quit R",
                 "yes, to save workspace image")
tbs_function(terminal, description)
```

Note:  [Link](https://www.r-bloggers.com/2013/03/download-and-install-r-in-ubuntu/) on how to download and install R through the terminal if not already downloaded to your machine. 

### 00_02_ github

From your github account create a repository nba_proj. 

```{r git, echo=FALSE}
terminal <- c("git init",
              "git remote add origin https://github.com/rbolt13/nba_proj.git",
              "git status",
              "git add *",
              "git status",
              "git add .Rhistory .Rprofile",
              "git status",
              "git commit -m \"my first commit\"",
              "git push origin main")
description <- c("initialize git repository",
                 "add location of repository",
                 "check status",
                 "select all",
                 "check status",
                 "add dot files that were missed",
                 "check status",
                 "adds comment to first commit",
                 "push to github")
tbs_function(terminal, description)
```

[Rename master branch to main](https://www.git-tower.com/learn/git/faq/git-rename-master-to-main)

# 01_ project_files.R, Makefile, and report.Rmd

<center>

<iframe src="https://giphy.com/embed/SyJyPwlThHScE" width="480" height="480" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p><a href="https://giphy.com/gifs/steven-universe-su-ruby-SyJyPwlThHScE">via GIPHY</a></p>

</center>

### 01_00_ scrape files 

Create the following files: 

* [00_00_scrape_stats_raw_data.R](https://github.com/rbolt13/nba_proj/blob/master/00_R/00_00_scrape_stats_raw_data.R)

* [00_01_scrape_colors_raw_data.R](https://github.com/rbolt13/nba_proj/blob/master/00_R/00_01_scrape_colors_raw_data.R)

Set environmental variable: 

```{r, echo=FALSE}
terminal <- c("export which_team=\"blazers\"",
              "R",
              "Sys.getenv(\"which_team\")",
              "q()",
              "y")
description <- c("create global variable",
                 "open R",
                 "verify that which_team = \"blazers\"",
                 "quit R",
                 "yes")
tbs_function(terminal, description)
```

Create lines 1-7 of the [Makefile](https://github.com/rbolt13/nba_proj/blob/main/Makefile), and save the raw data for stats and colors. 

```{r scrape data makefile, echo=FALSE}
terminal <- c("make raw_data stats",
              "make raw_data/colors")
description <- c("creates stats data file in raw data folder",
                "creates colors data file in raw data folder")
tbs_function(terminal, description)
```

**Push files to github.**

### 01_02_ label data

Create the following file: 

* [00_02_graph_labels_raw_data.R](https://github.com/rbolt13/nba_proj/blob/main/00_R/00_02_labels_raw_data.R)

Add lines 9-10 of the [Makefile](https://github.com/rbolt13/nba_proj/blob/main/Makefile), and save the label data. 
```{r labels file, echo=FALSE}
terminal <- c("make raw_data/labels")
description <- c("creates labels data file in raw data folder")
tbs_function(terminal, description)
```

**Push files to github.**

### 01_03_ clean data 

Create the following file: 

* [01_00_clean_data.R](https://github.com/rbolt13/nba_proj/blob/main/00_R/01_00_clean_data.R)

Set environmental variable: 

```{r, echo=FALSE}
terminal <- c("export which_stat=\"PTS\"",
              "R",
              "Sys.getenv(\"which_stat\")",
              "q()",
              "y")
description <- c("create global variable",
                 "open R",
                 "verify that which_stat = \"PTS\"",
                 "quit R",
                 "yes")
tbs_function(terminal, description)
```

Add lines 13-14 of the [Makefile](https://github.com/rbolt13/nba_proj/blob/main/Makefile), and save the clean data. 

```{r clean data makefile, echo=FALSE}
terminal <- c("make clean_data/raw_data")
description <- c("make clean data out of raw data")
tbs_function(terminal, description)
```

**Push files to github.**

### 01_05_ report

Create the following file: 

* [report.Rmd](https://github.com/rbolt13/nba_proj/blob/main/00_R/report.Rmd)

Add lines 22 - 23 of the [Makefile](https://github.com/rbolt13/nba_proj/blob/main/Makefile), and save report to output file. 

```{r report makefile, echo=FALSE}
terminal <- c("report")
description <- c("save report to output folder")
tbs_function(terminal, description)
```

**Push files to github.**

### 01_06_ env (package management)

Add lines 26 - 27 of the [Makefile](https://github.com/rbolt13/nba_proj/blob/main/Makefile), and update project. 

```{r make env file, echo=FALSE}
terminal <- c("restore")
description <- c("opens R and restores packages")
tbs_function(terminal, description)
```

**Push files to github.**

# 02_ Dockerfile

<center>

<iframe src="https://giphy.com/embed/VeNMIIutNd12576kaM" width="480" height="270" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p><a href="https://giphy.com/gifs/cartoonnetwork-steven-universe-the-movie-sutm-VeNMIIutNd12576kaM">via GIPHY</a></p>

</center>

Open docker on your computer. 

```{r, echo=FALSE}
terminal <- c("docker build -t nba_proj .",
              "docker run -it nba_proj /bin/bash")
description <- c("build docker",
                 "run docker interactive nba_proj docker with bash terminal")
tbs_function(terminal, description)
```

To be continued...

# Bonus Notes

* Makefile uses tabs not spaces, and you will get an error about missing separator. 

* git add --all 