---
title: ggplot2 UO Workshop
author: Randi Bolt
date: '2022-01-18'
slug: []
categories:
  - undergrad
tags:
  - r studio
  - undergrad
  - package
  - ggplot2
---

These are my notes from the University of Oregon's Data Service Consultants workshop on ggplot2. 

Gabriele Hayden with help from Cameron Mulder ran the workshop. They can be contacted via their consultation page [here](https://library.uoregon.edu/research-data-management/consultations). 

```{r setup}
# install.pcakages("ggpot2")
library(ggplot2)
```

```{r data}
download.file("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder_data.csv", destfile = "gapminder_data.csv")
gapminder <- read.csv("gapminder_data.csv", stringsAsFactors = TRUE)
```

```{r first example }
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) + 
  geom_point()
```

## Challenge 1

Modify the example so the figure shows how life expectancy has changed over time 

```{r}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp)) + 
  geom_point()
```

At this point my notes didn't save so there is a gap here between challenge 1 and 4. 

```{r}
americas <- gapminder[gapminder$continent == "america",]
lifeExp_plot <- ggplot(americas, aes(x = year, y = LifeExp)) + 
  geom_line() +
  facet_wrap(~ country) +
  labs(
    x = "Year",
    y = "Life expectancy",
    title = "Figure 1",
    color = "Continent"
  ) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
lifeExp_plot

# ggsave(filename = "LifeExp.png", plot = lifeExp_plot, width = 12, height = 10, dpi = 300, units = "cm")
```

## Challenge 4

Generate boxplots to compare the life expectancy between the different continents during the available years

```{r challenge 4}
ggplot(gapminder, aes(x=year, y=lifeExp, fill = continent)) +
  geom_boxplot() +
  labs(
    x = "Continent",
    y = "Life Expectancy",
    title = "Figure 1",
    fill = "Continent"
  )
```

```{r by year}
ggplot(data = gapminder, mapping = aes(x = continent, y = lifeExp, fill = continent)) +
  geom_boxplot() +
  facet_wrap( ~ year)+ 
  theme(axis.text.x = element_text(angle = 45))
```

```{r Dante's code}
ggplot(gapminder, aes(x = continent, y = lifeExp)) +
  geom_point(show.legend = FALSE, size = 0.75) +
  geom_boxplot(aes(fill = continent), alpha = 0.5) +
  facet_wrap(~ year) +
  labs(
    x = "Continent", #Title of x axis
    y = "Life Expectancy (years)", #title of y axis
    title = "Figure 2", #figure title
    color = "Continent" #title of legend
  ) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Thurseday 

- ggplot cookbook examples 

- integrate map data 

- plotly 


