---
title: Tidyverse Style Guide (Part 1)
author: Randi Bolt
date: '2022-03-29'
slug: []
categories: 
  - undergrad
tags: []
output:
  blogdown::html_page:
    toc: TRUE
---

```{r, echo=F, message=FALSE, error=FALSE, warning=FALSE}
x <- 1
y <- 2
library(datasets)
library(tidyverse)
data(iris)
```

This is part 1 of my notes from [the tidyverse style guide](https://style.tidyverse.org/).

# Analyses

## 1. Files 

- names are meaningful and use letters, numbers, `_`, and/or `-`. 

- avoid special characters in files names. 

- if more than 10 files then left pad with a zero (i.e. 00, 01, 02, 03,...)

- rename files instead of attempting 02a, 02b, and so on.  

- pay attention to capitalization. 

- use `-` and `=` to break up file into readable chunks. 

- load all packages at once in the beginning of the file. 

```{r}
# Load data ---------------------------

# Plot data ---------------------------
```

## 2. Syntax 

### Object names

- Variables and function names should use lovercase letters, numbers and `_` (no camel case).

- Base R uses dots in function and class names (`data.frame`).

- Variable names should be names and nouns. 

- Function names should be verbs. 

- Names should be concise and meaningful. 

- Avoid re-using names of common function and variables. 

### Spacing 

- Put a space after a common, never before. 

- Do not put spaces inside or outside parentheses for regular functions : `mean(x, na.rm = TRUE)`. 

- Put a space before and after () when using if, for, or while.

```{r}
# if (debug) {
#   show(x)
# }
```


- Place a space after `()` used for function arguments : `function(x) {}`.

- The embracing operator, `{{ }}`, should always have inner spaces to help emphasize its special behavior.

- Most infix operators (`==`, `+`, `-`, `<-`, etc.) should always be surrounded by spaces.

- Exceptions being: 

  - The operators with high precedence: `::`, `:::`, `$`, `@`, `[`, `[[`, `^`, `unary -`, `unary +`, and `:`.
  
  - Single-sided formulas when the right-hand side is a single identifier: `call(!!xyz)`.
  
  - When used in tidy evaluation `!!` (bang-bang) and `!!!` (bang-bang-bang) (because have precedence equivalent to unary `-/+`).
  
  - The helper operator: `package?stats`.
  
- adding extra spaces is ok if it improves alignment of `=` or `<-`.

### Function Calls 

- When you call a function omit the names of data arguemnts (such as `x=` because it is used so commonly).

- If you override the default value of an argument, use the full name.

- Avoid assignments in function calls.

  - Exception: function that capture side-effects: `output <- capture.output(x <- f())`.

### Control flow

- For `{}` brackets: 

  - `{` should be the last character on the line. 
  
  - Contents should be indented by two spaces. 
  
  - `}` should be the first character on the line. 
  
  - Else should be on the same line as `}` (if used).
  
- `{}` define the most important hierarchy of R code.

- Always use `&&` and `||` inside an if clause and never `&` and `|`. 

- If you want to rewrite a simple but lengthy if block:

```{r}
if (x > 10) {
  message <- "big"
} else {
  message <- "small"
}
```

Just write it all on one line:

```{r}
message <- if (x > 10) "big" else "small"
```

- it is okay to drop curly braces for very simple statements that fit on one line as long as they dont have side-effects.

- Function calls that affect control flow (like return(), stop() or continue) should always go in their own {} block.

- Avoid implicit type coercion (e.g. from numeric to logical) in if statements:

```{r}
# Good
if (length(x) > 0) {
  # do something
}

# Bad
if (length(x)) {
  # do something
}
```

- Avoid postion-based `switch()` statements. 

- Each element should go on its own line. 

- Provided a fall-through erro, unless you have previously validate the input. 

```{r}
switch(x, 
  a = 0,
  b = 1, 
  c = 2,
  stop("Unknown `x`", call. = FALSE)
)
```

### Long lines

- Limit code to 80 characters per line. This fits comfortably on a printed page with a reasonably sized font. 

```{r}
# # Good
# do_something_very_complicated(
#   something = "that",
#   requires = many,
#   arguments = "some of which may be long"
# )
# 
# # Bad
# do_something_very_complicated("that", requires, many, arguments,
#                               "some of which may be long"
#                               )
```

- You may also place several arguments on the same line if they are closely related to each other, e.g., strings in calls to paste() or stop().

### Semicolons

- Don’t put ; at the end of a line, and don’t use ; to put multiple commands on one line.

### Assignment

- Use <-, not =, for assignment.

### Data

- Use `"`, not `'`, for quoting text. The only exception is when the text already contains double quotes and no single quotes.

- Prefer TRUE and FALSE over T and F.

### Comments

- Each line of a comment should begin with the comment symbol and a single space.

- In data analysis code, use comments to record important findings and analysis decisions.

- If you need comments to explain what your code is doing, consider rewriting your code to be clearer. 

- If you discover that you have more comments than code, consider switching to R Markdown.

## 3. Functions 

### Naming 

- Strive to use verbs for function names.

### Long lines 

- Prefer function-indent style to double-indent style when it fits. 

```{r}
# Function-indent --------------------------------------
long_function_name <- function(a = "a long argument",
                               b = "another argument",
                               c = "another long argument") {
  # As usual code is indented by two spaces.
}

# Double-indent ----------------------------------------
long_function_name <- function(
    a = "a long argument",
    b = "another argument",
    c = "another long argument") {
  # As usual code is indented by two spaces.
}
```

- If a function argument can’t fit on a single line, this is a sign you should rework the argument to keep it short and sweet.

### return()

- Only use return() for early returns. Otherwise, rely on R to return the result of the last evaluated expression.

```{r}
# Good
find_abs <- function(x) {
  if (x > 0) {
    return(x)
  }
  x * -1
}
add_two <- function(x, y) {
  x + y
}

# Bad
add_two <- function(x, y) {
  return(x + y)
}
```

- Return statments should always be on their own line because they have important effects on the control flow. 

- If your function is called primarily for its side-effects (like printing, plotting, or saving to disk), it should return the first argument invisibly. This makes it possible to use the function as part of a pipe.

### Comments 

- In code, use comments to explain the “why” not the “what” or “how”.

- Comments should be in sentence case, and only end with a full stop if they contain at least two sentences. 

```{r}
# Good

# Objects like data frames are treated as leaves

# Do not use `is.list()`. Objects like data frames must be treated
# as leaves.

# Bad

# objects like data frames are treated as leaves

# Objects like data frames are treated as leaves.
```

## 4. Pipes 

### Introduction 

- Reserve pipes for a sequence of steps applied to one primary object.

- Avoid using pipes when there are meaningful intermediate objects that could be given informative names.

### Whitespace 

- `%>%` should always have a space before it, and should usually be followed by a new line. After the first step, each line should be indented by two spaces.

### Long lines

- If the arguments to a function don’t all fit on one line, put each argument on its own line and indent. 

### Short pipes

- A one-step pipe can stay on one line, but unless you plan to expand it later on, you should consider rewriting it to a regular function call.

- Sometimes it’s useful to include a short pipe as an argument to a function in a longer pipe. Carefully consider whether the code is more readable with a short inline pipe (which doesn’t require a lookup elsewhere) or if it’s better to move the code outside the pipe and give it an evocative name.

```{r, error=FALSE}
# Good
# x %>%
#   select(a, b, w) %>%
#   left_join(y %>% select(a, b, v), by = c("a", "b"))
# 
# # Better
# x_join <- x %>% select(a, b, w)
# y_join <- y %>% select(a, b, v)
# left_join(x_join, y_join, by = c("a", "b"))
```

### No arguments 

- magrittr allows you to omit () on functions that don’t have arguments. Avoid this feature.

### Assigment 

There are three acceptable forms of assignment:

- Variable name and assignment on separate lines:

```{r}
iris_long <-
  iris %>%
  gather(measure, value, -Species) %>%
  arrange(-value)
```

- Variable name and assignment on the same line:

```{r}
iris_long <- iris %>%
  gather(measure, value, -Species) %>%
  arrange(-value)
```

- Assignment at the end of the pipe with `->`:

```{r}
iris %>%
  gather(measure, value, -Species) %>%
  arrange(-value) ->
  iris_long
```

- The magrittr package provides the `%<>%` operator as a shortcut for modifying an object in place. Avoid this operator.

## 5. ggplot2

### Introduction 

- Styling suggestions for `+` used to separate ggplot2 layers are very similar to those for `%>%` in pipelines.

### Whitespace

- `+` should always have a space before it, and should be followed by a new line.

- If you are creating a ggplot off of a dplyr pipeline, there should only be one level of indentation.

### Long lines 

- If the arguments to a ggplot2 layer don’t all fit on one line, put each argument on its own line and indent. 

- ggplot2 allows you to do data manipulation, such as filtering or slicing, within the data argument. Avoid this, and instead do the data manipulation in a pipeline before starting plotting.


