---
title: 'Portland R User Group: Aggregate making tables with gt package'
author: Randi Bolt
date: '2022-02-24'
slug: []
categories:
  - undergrad
tags:
  - rstudio
---

Ted Landeras led this event for [Portland R user Grou](https://pdxrlang.com/)p where we watched [Tich Lannone|| Making Beautiful Tables with `{gt}`|| RStudio](https://www.youtube.com/watch?v=z0UGmMOxl-c) as a group on Youtube, and then went through some other examples. 

```{r}
library(gt)
library(tidyverse)
library(palmerpenguins)
```

use markdown with `md()`

pixel size `px()`

```{r}
penguins %>%
  group_by(species) %>%
  summarize_at(
    .vars = c(
      "bill_length_mm", 
      "bill_depth_mm",
      "flipper_length_mm",
      "body_mass_g"
      ),
            .funs = ~ mean(.,na.rm = TRUE)) %>%
  gt() %>%
  tab_header(
    title = md("The penguins dataset"),
    subtitle = md("**Three** years of data on penguins on _three_ islands.")
  ) %>%
  cols_label(
    bill_length_mm = md("Bill Length,<br>mm"),
    bill_depth_mm = md("Bill Depth,<br>mm"),
    flipper_length_mm = md("Flipper Length,<br>mm"),
    body_mass_g = md("Body Mass,<br>g")
  ) %>%
  opt_align_table_header(align = "left") %>% 
#   fmt_number(columns = everything()) %>%
#   fmt_number(columns = body_mass_g, scale_by = 1/1000) %>%
  cols_width(
    bill_length_mm ~ px(120),
    bill_depth_mm ~ px(120),
    flipper_length_mm ~ px(120),
    body_mass_g ~ px(120),
    everything() ~ px(100)
  ) %>% 
  tab_source_note(source_note = md("Dataset is from the `palmerpenguins` **R** package.")
                  ) %>%
  # tab_footnote(
  #   footnote = "The Largest of the three penguins studied.",
  #   locations = cells_stub(rows = "Gentoo")
  # )
  tab_footnote(
    footnote = "Flipper Length was measured with a tape measure.",
    locations = cells_column_labels(columns = flipper_length_mm)
  ) %>%
  tab_footnote(
    footnote = "Tap measure suffered some frost damage.",
    locations = list(
      cells_column_labels(columns = flipper_length_mm),
      cells_body(columns = flipper_length_mm)
  )
  ) %>%
  tab_style(
    locations = cells_body(
      columns = everything()
      ),
    style = list(
      cell_fill(color = "steelblue"),
      cell_text(color = "white")
      )
  ) %>%
  tab_stubhead(label = md("Penguin *Species*"))
```

# Practice 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remotes::install_github("jthomasmock/gtExtras")

#Main packages
library(gtsummary) 
library(gt) 
library(gtExtras)

#required for examples
library(survival)
library(dplyr) 
library(stringr) 
library(purrr) 
library(forcats)
library(tidyr)
```

# What can {gt} do for you?

The following are taken from a number of different examples. Where possible, I've tried to notate the source.


## {gt} parts reference

```{r}
url <- "gt_parts_of_a_table.svg"
knitr::include_graphics(url)
```





