---
title: 'Portland R User Group: Aggregate making tables with gt package'
author: Randi Bolt
date: '2022-02-24'
slug: []
categories:
  - undergrad
tags:
  - rstudio
---

Ted Landeras led this event for [Portland R user Group](https://pdxrlang.com/) where we watched [Tich Lannone|| Making Beautiful Tables with `{gt}`|| RStudio](https://www.youtube.com/watch?v=z0UGmMOxl-c) as a group on Youtube, and then went through some other examples. 

Additional Links: 

[Portland R User Group](https://pdxrlang.com/)

[Code of Conduct](https://pdxrlang.com/code-of-conduct.html)

```{r}
library(gt)
library(tidyverse)
library(palmerpenguins)
```

use markdown with `md()`

pixel size `px()`

```{r}
penguins %>%
  group_by(species) %>%
  summarize_at(
    .vars = c(
      "bill_length_mm", 
      "bill_depth_mm",
      "flipper_length_mm",
      "body_mass_g"
      ),
            .funs = ~ mean(.,na.rm = TRUE)) %>%
  gt() %>%
  tab_header(
    title = md("The penguins dataset"),
    subtitle = md("**Three** years of data on penguins on _three_ islands.")
  ) %>%
  cols_label(
    bill_length_mm = md("Bill Length,<br>mm"),
    bill_depth_mm = md("Bill Depth,<br>mm"),
    flipper_length_mm = md("Flipper Length,<br>mm"),
    body_mass_g = md("Body Mass,<br>g")
  ) %>%
  opt_align_table_header(align = "left") %>% 
  fmt_number(columns = bill_length_mm, decimals = 2) %>%
  fmt_number(columns = bill_depth_mm, decimals = 2) %>%
  fmt_number(columns = flipper_length_mm, decimals = 2) %>%
  fmt_number(columns = body_mass_g, scale_by = 1/1000) %>%
  cols_width(
    bill_length_mm ~ px(120),
    bill_depth_mm ~ px(120),
    flipper_length_mm ~ px(120),
    body_mass_g ~ px(120),
    everything() ~ px(100)
  ) %>% 
  tab_source_note(source_note = md("Dataset is from the `palmerpenguins` **R** package.")
                  ) %>%
  # tab_footnote(
  #   footnote = "The Largest of the three penguins studied.",
  #   locations = cells_stub(rows = "Gentoo")
  # )
  tab_footnote(
    footnote = "Flipper Length was measured with a tape measure.",
    locations = cells_column_labels(columns = flipper_length_mm)
  ) %>%
  tab_footnote(
    footnote = "Tap measure suffered some frost damage.",
    locations = list(
      cells_column_labels(columns = flipper_length_mm),
      cells_body(columns = flipper_length_mm)
  )
  ) %>%
  # tab_style(
  #   locations = cells_body(
  #     columns = c("bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g")
  #     ),
  #   style = list(
  #     cell_fill(color = "steelblue"),
  #     cell_text(color = "white")
  #     )
  # ) %>%
  data_color(
    columns = c(bill_depth_mm),
    colors = scales::col_numeric(
      palett = c("red", "orange", "brown"),
      domain = c(10,20)) 
  ) %>%
    data_color(
    columns = c(bill_length_mm),
    colors = scales::col_numeric(
      palett = c("red", "orange", "green"),
      domain = c(30,50)) 
  ) %>%
  tab_stubhead(label = md("Penguin *Species*")) %>%
  opt_table_font(font = google_font("Montserrat"), 
                 weight = 600, 
                 style = "italic")
```

# Practice 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remotes::install_github("jthomasmock/gtExtras")

#Main packages
library(gtsummary) 
library(gt) 
library(gtExtras)

#required for examples
library(survival)
library(dplyr) 
library(stringr) 
library(purrr) 
library(forcats)
library(tidyr)
```

# What can {gt} do for you?

The following are taken from a number of different examples. Where possible, I've tried to notate the source.


## {gt} parts reference

```{r}
url <- "gt_parts_of_a_table.svg"
knitr::include_graphics(url)
```

## Allison Horst's Examples

These examples are from: https://www.allisonhorst.com/post/2020-03-02-gt-tables-examples/

### Example 1

```{r}
head(rock)
```


```{r}
rock %>% # Get 'rock' data
  head(5) %>% # First 5 lines only
  gt() # Make a table, it just works.
```
### Example 1

```{r}
head(BOD)
```

```{r}
BOD %>% # Get the data...
gt() %>% # use 'gt' to make an awesome table...
  tab_header( 
    title = "BOD Table Woooooo!", # ...with this title
    subtitle = "Hooray gt!") %>% # and this subtitle
  fmt_number( # A column (numeric data)
    columns = vars(Time), # What column variable? BOD$Time
    decimals = 2 # With two decimal places
    ) %>% 
  fmt_number( # Another column (also numeric data)
    columns = vars(demand), # What column variable? BOD$demand
    decimals = 1 # I want this column to have one decimal place
  ) %>% 
  cols_label(Time = "Time (hours)", demand = "Demand (mg/L)") # Update labels
```

## Example 2

```{r}
tooth_length <- ToothGrowth %>% 
  group_by(supp, dose) %>% 
  summarize(
    mean_len = mean(len)
  ) %>% 
  as_tibble() 

# A gt table: 
tooth_length %>% # Take tooth_length
  gt() %>% # Make a gt table with it
  tab_header(
    title = "A title just like that", # Add a title
    subtitle = "(with something below it!)" # And a subtitle
  ) %>%
  fmt_passthrough( # Not sure about this but it works...
    columns = vars(supp) # First column: supp (character)
  ) %>% 
  fmt_number(
    columns = vars(mean_len), # Second column: mean_len (numeric)
    decimals = 2 # With 4 decimal places
  ) %>% 
  fmt_number(
    columns = vars(dose), # Third column: dose (numeric)
    decimals = 2 # With 2 decimal places
  ) %>% 
  data_color( # Update cell colors...
    columns = vars(supp), # ...for supp column!
    colors = scales::col_factor( # <- bc it's a factor
      palette = c(
        "green","cyan"), # Two factor levels, two colors
      domain = c("OJ","VC")# Levels
  )
  ) %>% 
  data_color( # Update cell colors...
    columns = vars(dose), # ...for dose column 
    colors = scales::col_numeric( # <- bc it's numeric
      palette = c(
        "yellow","orange"), # A color scheme (gradient)
      domain = c(0.5,2) # Column scale endpoints
  )
  ) %>% 
  data_color( # Update cell colors...
    columns = vars(mean_len), # ...for mean_len column
    colors = scales::col_numeric(
      palette = c(
        "red", "purple"), # Overboard colors! 
      domain = c(7,27) # Column scale endpoints
  )
  ) %>% 
  cols_label(supp = "Supplement", dose = "Dosage (mg/d)", mean_len = "Mean Tooth Length") %>% # Make the column headers
  tab_footnote(
    footnote = "Baby footnote test", # This is the footnote text
    locations = cells_column_labels(
      columns = vars(supp) # Associated with column 'supp'
      )
    ) %>% 
    tab_footnote(
    footnote = "A second footnote", # Another line of footnote text
    locations = cells_column_labels( 
      columns = vars(dose) # Associated with column 'dose'
      )
    )
```

## Animal Crossing (by Ted)

This is an example using the Animal Crossing dataset from Tidy Tuesday. In this dataset, you'll see that the `items` column is a URL. 

```{r}
items <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/items.csv')
head(items)
```

We can display 

```{r}
items %>% 
  top_n(10, buy_value) %>%
  arrange(desc(buy_value)) %>%
  select(name, sell_value, buy_value, category, image=image_url) %>%
  gt() %>%
   text_transform(
     #tell gt that the image column 
    locations = cells_body(vars(image)),
    fn = function(x) {
      web_image(
        url = x,
        height = 50
      )
    }
  )

```

## Tom Mock's example with {gtExtras}

Tom Mock has created a package called `{gtExtras}` which can be installed with the following command:

```{r}
library(gtExtras)
remotes::install_github("jthomasmock/gtExtras")
```

This example is from one of Tom's tweets. Gist is here: https://gist.github.com/jthomasmock/923604deb65682b0364c9220a85ddb36

```{r}
stream_data <- readr::read_csv("stream_data.csv")

stream_data
```


```{r}
gt(stream_data) %>% 
  #transform Nominee and Winner columns into a bullet plot
  gt_plt_bullet(column = Nominee, target = Winner) %>% 
  #take the type column and retrieve a font awesome symbol
  gt_fa_column(column = type) %>% 
  #format table with nytimes style formating
  gt_theme_nytimes() %>% 
  #format ratio column with percentage
  fmt_symbol_first(column = Ratio, suffix = "%", decimals = 1, scale_by = 100)
```

## {gtsummary} package

Example from: https://cran.r-project.org/web/packages/gtsummary/vignettes/gallery.html

For biostatisticians, one of the most annoying tasks to build is a Table 1: a breakdown of research subjects by various demographics. The {gtsummary} package simplifies building these tables.

```{r}
trial %>%
  select(trt, age, grade) %>%
  tbl_summary(
    by = trt, 
    missing = "no",
    statistic = all_continuous() ~ "{median} ({p25}, {p75})"
  ) %>%
  modify_header(all_stat_cols() ~ "**{level}**<br>N =  {n} ({style_percent(p)}%)") %>%
  add_n() %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**Chemotherapy Treatment**")
```





