---
title: UO ggplot2 p.2
author: Randi Bolt
date: '2022-01-20'
slug: []
categories:
  - undergrad
tags:
  - ggplot2
  - package
  - r studio
  - undergrad
  - plotly
  - rcookbook
  - maps
  - sf
  - rmapshaper
  - flametree 
---

These are my notes from the University of Oregon's Data Service Consultants workshop on ggplot2, part 2. 

Cameron Mulder ran this workshop. He can be contacted via their consultation page [here](https://library.uoregon.edu/research-data-management/consultations). 

```{r set up, error=FALSE}
# library(ggplot2)
# install all of tidyverse 
# install.packages("tidyverse")
# loads entire tidyverse which inclused ggplot, dyply, and so on 
library(tidyverse)
```


## ggplot2 review 

`ggplot(data= <DATA>, mapping = aes(<MAPPING>))+ <GEOM FUNCTION>()`

```{r ggplot review}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point()
```

```{r scatterplot}
# base R
plot(mpg$displ, mpg$hwy)
```

Note : the `$` is how we id the specific variable we are wanting to work with. 

## Line Graph (Base R)

```{r line graph }
# help(pressure) , is the same as ?pressure
plot(pressure$temperature, pressure$pressure, type = "l")
# add points
points(pressure$temperature, pressure$pressure)
# add lines (and points)
lines(pressure$temperature, pressure$pressure/2, col = "red")
points(pressure$temperature, pressure$pressure/2, col = "red")
```


## Line Graph (ggplot)

```{r line graph ggplot}
ggplot(pressure, aes(x = temperature, y = pressure)) + 
  geom_line() + 
  geom_point()
```

## Bar Graphs (base R)

```{r bar graph base r}
# using the BOD dataset 
# Biochemical Oxygen Demand Dataset 
barplot(BOD$demand, names.arg = BOD$Time)
```

```{r}
# using mtcars to look at count 
# look at some stats 
table(mtcars$cyl)
# barplot 
barplot(table(mtcars$cyl))
```

## Bar Graphs (ggplot)

```{r bar graph BOD}
# need to change time into a factor() because otherwise it is continuous 
ggplot(BOD, aes(x = factor(Time), y = demand)) + 
  geom_col()
```

Notice that the 6 isn't there, that is because of factor(). 

Note : geom_bar does counts, but column has the height of the bar based on the data. 

```{r bargraph ggplot cars}
# don't need to id the y becasue the y will be the count.
ggplot(mtcars, aes(x=cyl)) +
  geom_bar()
```

## Histogram (Base R)

```{r base r histogram}
hist(mtcars$mpg)
hist(mtcars$mpg, breaks = 4)
```

## Histogram (ggplot)

```{r}
ggplot(mtcars, aes(x=mpg)) +
  geom_histogram(binwidth = 4)
```

## Boxplot (base r)

```{r}
# ?ToothGrowth
plot(ToothGrowth$supp, ToothGrowth$len)
# formula syntax
plot(len ~ supp, data = ToothGrowth)
plot(len ~ supp + dose, data = ToothGrowth)
```

## Challenge 1 

Create a boxplot of toothgrowth using len and supp from ToothGrowth data set. 

Hint: geom_boxplot

```{r challenge 1}
ggplot(ToothGrowth, aes(x= supp, y = len)) +
  geom_boxplot()
```

## Time Series 

ggplot will automatically recognize the variable as a date as long as the variable is imported as a date. 

```{r time series}
# dummy data
data <- data.frame(
  day = as.Date("2017-06-14")-0:364,
  value = runif(365)
)
head(data)
# plot 
p <- ggplot(data, aes(x = day, y = value)) +
  geom_line()
p
```

Now lets use the economics data set 

```{r}
# ?economics
ggplot(data = economics, aes(x = date, y = pop)) +
  geom_line()
# subset of data
subset <- economics %>%
  filter(date>as.Date("2006-1-1"))
ggplot(subset, aes(x = date, y = pop)) +
  geom_line()
# chart differet data over time
ggplot(economics, aes(x = date, y = pop)) +
  geom_line(aes(size = unemploy), color = "red")
```

The size of the line is based on the value of unemployment (which is the number of unemployment in thousands). 

```{r reshape data}
# psavert is the personal savings rate 
# unemployment median duration
econ_long <- economics %>% 
  select(date, psavert, uempmed) %>%
  pivot_longer(col = c('psavert', 'uempmed'),
               names_to = "variable",
               values_to = "values")
head(econ_long)
# data 
ggplot(data = econ_long, aes(x = date, y = values)) +
  geom_line(aes(color = variable), size = 1)
```

R is a great tool for plotting time series. 

## Maps!!

Polygon maps

```{r polygon maps}
mi_counties <- map_data("county", "michigan") %>%
  select(lon = long, lat, group, id = subregion)
head(mi_counties)
# map 1 
ggplot(mi_counties, aes(lon, lat))+
  geom_point() +
  coord_quickmap()
# nicer map (still not that common)
ggplot(mi_counties, aes(lon, lat, group = group))+
  geom_polygon(fill = "white", color = "grey") +
  coord_quickmap()
```

## Simple Features 

```{r}
# australlian maps
# install.packages('ozmaps')
library(ozmaps)
library(sf)
```

```{r}
# use :: to pull from specific packages
oz_stats <- ozmaps::ozmap_states
oz_stats
# map of australlia 
ggplot(oz_stats)+
  geom_sf() +
  coord_sf()
```

```{r rmapshaper}
# install.packages('rmapshaper')
library(rmapshaper)
```

```{r map}
oz_stats <- ozmaps::ozmap_states %>% filter(NAME != "Other Territories")
oz_votes <- rmapshaper::ms_simplify(ozmaps::abs_ced)
# map 
ggplot()+
  geom_sf(data = oz_stats, mapping = aes(fill = NAME)) +
  geom_sf(data = oz_votes, fill = NA) +
  coord_sf()
```

## Plotly

```{r plotly}
# install.packages("plotly")
library(plotly)
```


```{r interactive}
fig <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length)
fig
```

```{r plotly in tidyverse}
mpg %>% plot_ly(x = ~displ, y = ~mpg, color = ~class)
```

You can double click on the legend to see a subset of the data. 

```{r ggplot and plotly}
plot <- ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) +
  geom_smooth()
plot
ggplotly(plot)
```

## Art 

```{r art}
# install.packages("flametree")
library(flametree)
# pick some colors 
shades <- c("blue", "green", "red", "orange")
data <- flametree_grow(time = 10, trees = 10)
data %>% flametree_plot(
  background = "white",
  palette = shades,
  style = "nativeflora"
)
```

Package by Danielle Navarro. Check out her art [here](art.djnavarro.net). 

Next week we are going over rmd. 



